---
output:
  word_document: default
  html_document: default
---
Emilee Draper

Module 6 Clustering Assignment

```{r}
library(tidyverse)
library(cluster)
#install.packages("factoextra")
#install.packages("dendextend")
library(factoextra)
library(dendextend)
```

```{r}
trucks <- read_csv("trucks.csv")
View(trucks)
```

TASK 1
```{r}
ggplot(trucks, aes(x = Distance, y = Speeding)) +geom_point()
```
Yes, there seems to be natural clustering around distance of 25 - 75 with speedings between 0 - 18, and speedings between 20 -60. There is also clustering around distnace 150 - 220 with speedings between 0 - 25, and more loosely around speedings between 40 - 100. 

TASK 2
```{r}
trucks = trucks %>% select(-Driver_ID)
trucks2 = as.data.frame(scale(trucks))
str(trucks2)
```

TASK 3
```{r}
set.seed(1234)
clusters1 = kmeans(trucks2, 2)
fviz_cluster(clusters1, trucks2)
```

TASK 4

```{r}
set.seed(123)
fviz_nbclust(trucks2, kmeans, method ="wss")
```
```{r}
set.seed(123)
fviz_nbclust(trucks2, kmeans, method = "silhouette")
```
Both methods seem to agree on 4 as the appropriate number of clsuters. 

TASK 5
```{r}
set.seed(1234)
clusters2 <- kmeans(trucks2, 4)
fviz_cluster(clusters2, trucks2)
```

TASK 6

Cluster 1 is full of drivers who travel further, but have a lower percentage of speeding. Cluster 2 is full of dirivers who travel shorter distance, and have a lower percentage of speeding. Cluster 3 is full of drivers who travel further distance, and have a medium and higher percentage of speeding. Cluster 4 is full of drivers who travel shorter, but have a  medium percentage of speeding. 


```{r}
wine <- read_csv("wineprice.csv")
View(wine)
```

```{r}
wine = wine %>% select(-Year)
wine = wine %>% select(-FrancePop)
wine2 = as.data.frame(scale(wine))
str(wine2)
```

TASK 7
```{r}
set.seed(123)
fviz_nbclust(wine2, kmeans, method ="wss")
```

```{r}
set.seed(123)
fviz_nbclust(wine2, kmeans, method = "silhouette")
```
The first graph I would think 5 or 6 but the second graph shows 5, so I am going to choose 5 as the optimal amount of clusters.

TASK 8
```{r}
set.seed(1234)
clusters <- kmeans(wine2, 5)
fviz_cluster(clusters, wine2)
```

TASK 9
```{r}
m = c("average", "single", "complete", "ward")
names(m) = c("average", "single", "complete", "ward")

ac = function(x) {
  agnes(wine2, method = x)$ac
}
map_dbl(m, ac)

```
```{r}
hc = agnes(wine2, method = "ward")
pltree(hc,cex = 0.6, hang = -1, main = "Agglomerative Dendrogram")

plot(hc, cex.axis = .5)
rect.hclust(hc, k = 5, border =2:6)
```

TASK 10
```{r}
hc2 = diana(wine2)
pltree(hc2, cex = 0.6, hang = -1, main = "Divisive Dendogram")

plot(hc2, cex.axis = .5)
rect.hclust(hc2, k = 5, border =2:6)
```

