---
title: "Module 5 Assignment 1"
output: word_document
---
```{r}
library(tidyverse)
library(caret)
#install.packages("nnet")
library(nnet)
```

```{r}
library(readr)
parole <- read_csv("~/Desktop/BAN 502/module 4/parole.csv")
View(parole)
```

```{r}
parole = parole %>% mutate(male = as.factor(as.character(male))) %>%
	mutate(male = fct_recode(male, "male" = "0" , "female" = "1")) 
  
parole = parole %>% mutate(race = as.factor(as.character(race))) %>% 
  mutate(race = fct_recode(race, "white" = "1", "otherwise" = "2"))

parole = parole %>% mutate(state = as.factor(as.character(state))) %>%
  mutate(state = fct_recode(state, "kentucky" = "2", "Louisiana" = "3", "Virginia" = "4", "other" = "1")) 

parole = parole %>% mutate(crime = as.factor(as.character(crime))) %>%
  mutate(crime = fct_recode(crime, "larceny" = "2", "drug-related" = "3", "driving-related" = "4", "other" = "1"))

parole = parole %>% mutate(multiple.offenses = as.factor(as.character(multiple.offenses))) %>%
	mutate(multiple.offenses = fct_recode(multiple.offenses, "Yes" = "1", "No" = "0")) 

parole = parole %>% mutate(violator = as.factor(as.character(violator))) %>%
	mutate(violator = fct_recode(violator, "Yes" = "1", "No" = "0"))
	str(parole)
```

TASK 1
```{r}
set.seed(12345)
train.rows = createDataPartition(y=parole$violator, p=.7, list = FALSE)	
train = parole[train.rows,]
test = parole[-train.rows,]
```

TASK 2
```{r}
start_time = Sys.time()

fitControl = trainControl(method = "cv",			
				number = 10)

nnetGrid <- expand.grid(size = 12, decay = 0.1)	#12 because 6 variables but some have multiple levels, 0.1 is normal for decay 

set.seed(1234)
nnetBasic = train(violator ~ .,
		parole,
		method = "nnet",
		tunGrid = nnetGrid,
		trControl = fitControl,
		verbose = FALSE,
		trace = FALSE)

end_time = Sys.time()
end_time-start_time		

```

TASK 3
```{r}
nnetBasic
```

```{r}
predNetBasic = predict(nnetBasic, train)
```

```{r}
confusionMatrix(predNetBasic, train$violator, positive = "Yes")
```
The accuracy is rounds to 90% and naive accuracy is 88%. I would say the model is okay because it is indeed higher than the naive accuracy, but not by much. The model size used was 1 and the decay rate was .1

TASK 4
```{r}
Start_time = Sys.time()	
fitControl = trainControl(method = "cv",			
				number = 10)

nnetGrid <- expand.grid(size = seq(from = 1, to = 12, by = 1), 
                      decay = seq(from = 0.1, to = 0.5, by = 0.1))               
	 
set.seed(1234)
nnetFit = train(violator ~ .,
		parole,
		method = "nnet",
		tuneGrid = nnetGrid,
		trControl = fitControl,
		verbose = FALSE,
		trace = FALSE)

end_time = Sys.time()
end_time-start_time		#shows how long

```

TASK 5

```{r}
nnetFit
```
```{r}
plot(nnetFit)
```

```{r}
predNet = predict(nnetFit, train)

confusionMatrix(predNet, train$violator, positive = "Yes")
```
Accuracy and naive accuracy are the same as the model in Task 2. The model size used was 2 and the decay was .4, which are both larger than model in Task 2. 

TASK 6
```{r}
predNetBasic = predict(nnetBasic, test)
```

```{r}
confusionMatrix(predNetBasic, test$violator, positive = "Yes")
```
The accuracy is relatively the same for the training data. Sensitivitity is slighlty decreased and specificity is slightly increased compared to the training set prediction. 

TASK 7
```{r}
predNet = predict(nnetFit, test)

confusionMatrix(predNet, test$violator, positive = "Yes")
```
The accuracy is also the same for this testing set as the training set from Task 4. Sensitivity is slighltly decreased and specificity is slighlty increased in this prediction. 

TASK 8
The testing and training data perform almost exactly the same on both models, which leads me to believe that no overfitting has occured. 
